---
output:
  word_document: default
  html_document: default
---

#First, load packages
```{r}
library(utils)
library(base)
library(stats)
library(graphics)
library(cluster)
library(factoextra)
library(knitr)
```

#Second, prep the data
```{r}
#Load dataset
Cereal.Data<-read.csv("Cereals.csv") 

#Set row names to the Cereals column
row.names(Cereal.Data) <- Cereal.Data[,1]

#Remove Cereals column
Cereal.Data <- Cereal.Data[,-1]
```

#Normalize dataset
```{r}
Cereal.DF<-Cereal.Data[,3:15] #only numeric variables
Cereal.DF.Norm1<-sapply(Cereal.DF, scale) # normalization

row.names(Cereal.DF.Norm1) <- row.names(Cereal.Data)#set row names to cereal names
```

#Remove all cereals with missing values
```{r}
Cereal.DF.Norm2<-na.omit(Cereal.DF.Norm1)
head(Cereal.DF.Norm2)
```
#Dissimilarity matrix 
```{r}
d<-dist(Cereal.DF.Norm2, method = "euclidean")
```

#Hierarchical clustering using different methods
```{r}
hclust.1<-hclust(d, method = "single")
plot(hclust.1, cex=0.3, hang=-1, sub="single method")

hclust.2<-hclust(d, method="complete")
plot(hclust.2, cex=0.3, hang=-1, sub="complete method")

hclust.3<-hclust(d, method="average")
plot(hclust.3, cex=0.3, hang=-1, sub="average method")

hclust.4<-hclust(d, method="ward.D")
plot(hclust.4, cex=0.3, hang=-1, sub="ward method")
```

#Computing AGNES with different methods of linkage
```{r}
hclust.single<-agnes(Cereal.DF.Norm2, method="single")
hclust.complete<-agnes(Cereal.DF.Norm2, method="complete")
hclust.average<-agnes(Cereal.DF.Norm2, method= "average")
hclust.ward<-agnes(Cereal.DF.Norm2, method= "ward")
```

#Compare AGNES coefficients with different linkage methods
```{r}
print(hclust.single$ac)
print(hclust.complete$ac)
print(hclust.average$ac)
print(hclust.ward$ac) #Best method of linkage
```
#The best method of linkage is the ward method. This is because the AGNES coefficient is the highest;therefore, the ward method provides the strongest clustering structure.

#How many clusters would you choose? 4 clusters
```{r}
plot(hclust.4,cex=0.3, hang=-1)
rect.hclust(hclust.4, k=4, border="purple")
clust.1<-cutree(hclust.4, k=4)
c.1<-as.data.frame(cbind(Cereal.DF.Norm2,clust.1))
```

#Comment on the structure of the clusters and on their stability.
```{r}
#normalized data with missing values omitted
part.data<-Cereal.DF.Norm2

#partition the cereal data
part.A<-part.data[1:55,]
part.B<-part.data[56:74,]

#Computing AGNES with different methods of linkage
hclust.5<- agnes(part.A,method = "ward")
hclust.6<-agnes(part.A,method="average")
hclust.7<-agnes(part.A,method="complete")
hclust.8<-agnes(part.A,method="single")

#create a table of AGNES coefficients
kable(cbind(ward=hclust.5$ac,average=hclust.6$ac,complete=hclust.7$ac,single=hclust.8$ac))
#Ward method is the best because it has the highest AGNES coefficient. 

#plot dendrogram of AGNES using best method (ward method)
pltree(hclust.5,cex=0.3,hang=-1,main="Dendrogram of AGNES")

#create clusters in dendrogram of AGNES
rect.hclust(hclust.5, k = 4, border = "purple")

#identify cereals in the 4 clusters
clust.2<-cutree(hclust.5, k=4)
result<-as.data.frame(cbind(part.A,clust.2))

#finding cluster centers of part.A data
result[result$clust.2==1,]
center.1<-colMeans(result[result$clust.2==1,])
result[result$clust.2==2,]
center.2<-colMeans(result[result$clust.2==2,])
result[result$clust.2==3,]
center.3<-colMeans(result[result$clust.2==3,])
result[result$clust.2==4,]
center.4<-colMeans(result[result$clust.2==4,])
all.centers<-rbind(center.1,center.2,center.3,center.4)

#Assigning part.B data to clusters using cluster centroids from part.A data
c.2<-as.data.frame(rbind(all.centers[,-14],part.B))

dis<-get_dist(c.2)
matrix<-as.matrix(dis)

data.1<-data.frame(data=seq(1,nrow(part.B),1),clusters=rep(0,nrow(part.B)))

#Assessing how consistent the cluster assignments are compared to the assignments based on all data.
for(i in 1:nrow(part.B))
{
  data.1[i,2]<-which.min(matrix[i+4,1:4])
}
data.1
cbind(c.1$clust.1[56:74],data.1$clusters)
table(c.1$clust.1[56:74]==data.1$clusters)
```
#The stability of the clusters is 94.74% rounded. Out of the 19 variables in part.B, 18 were consistent with assignments based on all data and 1 was not consistent. Therefore, the stability is 18/19 = 94.74% rounded.

#Find a cluster of "healthy cereals" 
```{r}
#We do not want to normalize the data because we want the exact value of the ratings to determine the healthy cereals. We still want to omit missing values.
cereals<-na.omit(Cereal.Data)

#show data frame for each cluster to analyze ratings
cluster.1<-as.data.frame(cbind(cereals,clust.1))
cluster.1[cluster.1$clust.1==1,]
cluster.1[cluster.1$clust.1==2,]
cluster.1[cluster.1$clust.1==3,]
cluster.1[cluster.1$clust.1==4,]

```
#Based on the ratings of the various cereals in each cluster, cluster 1 consists of the highest rated cereals; therefore, cluster 1 has the healthiest cereals followed by cluster 4. Cluster 3 has the unhealthiest cereals due to the lowest ratings of cereals in that cluster. 



